A Type System for Safe and Scalable Parallelism using Logical Regions

### 2012-07-11 doc/popl2013/intro.tex
intro

In the past decade machine architecture, particularly at the high
performance end of the spectrum, has undergone a revolution.  The
latest supercomputers are now complex hierarchies of many different
kinds of computing technologies: networks of nodes at the top level,
multiple chips per node, multiple cores within a chip, and, most
recently, multiple accelerators (usually GPUs) per node, which 
can themselves be decomposed.  We present the operational and static
semantics of Legion\cite{Legion12}, a programming model targeted at providing an
appropriate level of abstraction for programming such machines, one
that is both sufficiently high-level to be portable while still
exposing aspects that are crucial to performance.

In Legion data is organized in a hierarchy of {\em regions}
and subregions while computation is organized in a hierarchy of {\em
tasks} and subtasks operating on regions.  Regions and tasks interact
through a static system of {\em region privileges} specifying 
operations a task may perform on a region argument (read,
write, or reduce) and  {\em region coherence} annotations that
express what other tasks may do concurrently with
the region (exclusive, atomic, or simultaneous).  We prove the
soundness of the privileges and coherence system and use these two
results to prove a third result: if two {\em siblings} (tasks that
have the same immediate parent task) $t_1$ and $t_2$ are 
{\em non-interfering} (have no ordering requirements for correctness), 
then the any descendant of $t_1$ in the task hierarchy is non-interfering
with any descendant of $t_2$.  This theorem is the basis for correct distributed
task scheduling in a Legion implementation, which is crucial for
performance; a centralized scheduler would be a bottleneck
because of the huge latencies involved in communicating in the target
class of machines.

>>>>

In the past decade machine architecture, particularly at the high
performance end of the spectrum, has undergone a revolution.  The
latest supercomputers are now complex hierarchies of many different
kinds of computing technologies: networks of nodes at the top level,
multiple chips per node, multiple cores within a chip, and, 
recently, multiple accelerators (usually GPUs) per node, which 
can themselves be further decomposed.  We present the operational and static
semantics of Legion\cite{Legion12}, a programming model targeted at providing an
appropriate level of abstraction for programming such machines, one
that is both sufficiently high-level to be portable while still
exposing aspects that are crucial to performance.

In Legion data is organized in a hierarchy of {\em regions}
and subregions while computation is organized in a hierarchy of {\em
tasks} and subtasks operating on regions.  Regions and tasks interact
through a static system of {\em region privileges} specifying 
operations a task may perform on a region argument (read,
write, or reduce) and  {\em region coherence} annotations that
express what other tasks may do concurrently with
the region (exclusive, atomic, or simultaneous).  We prove the
soundness of the privileges and coherence system and use these two
results to prove a third result: if two {\em siblings} (tasks that
have the same immediate parent task) $t_1$ and $t_2$ are 
{\em non-interfering} (have no ordering requirements for correctness), 
then the any descendant of $t_1$ in the task hierarchy is non-interfering
with any descendant of $t_2$.  This theorem is the basis for correct distributed
task scheduling in a Legion implementation, which is crucial for
performance; a centralized scheduler would be a bottleneck
because of the large communication latencies in the target
class of machines.
